---
title: "MachineLearningProject"
output: html_document
---

# Practical Machine Learning Project

## Data
### Load the training and testing data sets
```{r}
train <- read.csv( "/Users/golinger/DataScience/Machine Learning/Project/pml-training.csv" )
test <- read.csv( "/Users/golinger/DataScience/Machine Learning/Project/pml-testing.csv" )
```
### Look at the train data
```{r}
dim( train )
table( train$classe )
```
### Look at the test data
```{r}
dim( test )
table( test$classe )
```

# Preprocessing
## Partitioning
### Create a validation set
```{r}
library(caret)
set.seed(22222)
trainIx <- createDataPartition( train$classe, p = 0.8, list = FALSE )
Training <- train[ trainIx, ]
Validation <- train[ -trainIx, ]
```

## Feature Selection
### Remove near zero variance features
```{r}
nzvcol <- nearZeroVar( Training )
Training <- Training[ , -nzvcol ]
```

### Remove features with a lot of missing values
```{r}
cntLength <- sapply( Training, function(x) 
  {
  sum( !(is.na(x) | x == "" ))
  })
nullCol <- names( cntLength[cntLength < 0.4 * length( Training$classe )] )
```

### Remove features that are just descriptive
```{r}
descriptCol <- c( "X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
                 "cvtd_timestamp", "new_window", "num_window" )
excludeCols <- c( descriptCol, nullCol )
Training <- Training[, !names( Training ) %in% excludeCols]
```

# Model
## Model Training
```{r}
library( randomForest )
rfModel <- randomForest( classe ~ ., data = Training, importance = TRUE, ntrees = 10 )
```
## Model Validation
## Train Set Accuracy
```{r}
ptraining <- predict( rfModel, Training )
print( confusionMatrix( ptraining, Training$classe ))
```

## Validation set Accuracy
### This set is used to cross validate the model.
```{r}
pvalidation <- predict( rfModel, Validation )
print( confusionMatrix( pvalidation, Validation$classe ))
```
### The out of sample error is very small and well within any over fitting concerns.

# Test Set Prediction
```{r}
ptest <- predict( rfModel, test )
print( ptest )
```

## Test Set Files
```{r}
testSet <- as.vector( ptest )

writeTestFiles = function( x ) 
  {
  n = length(x)
  for ( i in 1:n ) 
    {
    filename = paste0( "problem_id_", i, ".txt" )
    write.table( x[i], file = filename, quote = FALSE, row.names = FALSE, 
                col.names = FALSE )
    }
  }

writeTestFiles( testSet )
```

